Search-based Synthesis of Network Intrusion Detection Rules

### 2019-09-14 paper/paper.tex
pass

\tname{} is a search-based algorithm that uses positive and negative
examples to synthesize signature-based rules, such as those from Snort
and Suricata. \tname{} synthesizes rules in two steps. First, it
produces an overspecified rule to capture the negative example
provided as input. Then, it uses positive examples to eliminate
unecessary options from the rule. The goal is to create a rule that
captures only the negative traffic.

@@

@@

>>>>

\tname{} is a search-based approach that uses positive and negative
network traffic examples to synthesize signature-based NIDS rules,
such as those from Snort and Suricata. \tname{} synthesizes rules in
two steps. First, it produces a potentially overspecified rule that
captures the negative example that is provided as input. Then, it uses
positive examples to eliminate unecessary options from the rule. The
goal is to create a rule that captures only the negative traffic.

@@

Search-based algorithms need to characterize the problem (\ie{}, a
solution candidate) in some way. In this case, the problem is the
rule.  \tname\ encodes the rule options segment (see
Section~\ref{sec:example-suricata-rules}) of a Suricata rule as a
vector of Option objects. For example, \tname{} models the option
\CodeIn{flags} from the example on Figure~\ref{fig:synflood-example}
with a data structure storing two fields, one of character type and
another of integer type. This enables \tname{} to mutate each element
separately.\Mar{Lucas, por favor, confirme.}

%% \tname{}
%% uses an abstract representation of a Suricata rule that reflects the
%% concrete representation of the rule. The concrete rule representation
%% is the one actually used by Suricata whereas the abstract
%% representation is the one \tname{} uses to run the search.

@@

The fitness value of a candidate solution $c$, modelling the rule
options $o_1, \dots, o_n$ is given by $f(c)=1-\Sigma{d(o_i,m)}/n$,
where $i$ ranges from 1 to $n$ and $m$ denotes the negative
traffic. The higher the fitness value the better, \ie{}, the fitter
the candidate solution $c$ will be.  The value of the expression
$d(o_i,m)$, associated with option $o_i$, ranges over
0-1. Consequently, the value of the expression $\Sigma{d(o_i,m)}/n$
itself is in the 0-1 range, as there are $n$ options, and the range of
the function is 0-1.

As usual, we used customized distance functions to compute $d(o_i,
m)$. For example, let us consider the \CodeIn{flags} option, discussed
on Figure~\ref{fig:synflood-example}. Recall that this option
identifies the SYN packet in the tcp protocol and that its encoding in
\tname{} stores two attributes---one categorical and one
numerical. The value of $d(o_i, m)$, in this case, is obtained by
computing the normalized Euclidian distance in the two-dimensional
space. Considering that each of the two dimensions has cardinality of
ten, the distance of a candidate solution holding the option
\CodeIn{flags(S,13)} to the solution will be 0.22
(=$1/\sqrt{20}$). Note that the smaller the distance the better.

\tname{} uses a modified version of Suricata to measure distance. It
updates the rule set that the modified version of Suricata analyses,
injects the traffic on the network, and monitors the execution. Each
rule option has an associated \CodeIn{eval} function in code. \tname{}
monitors the execution of these functions to identify how far it is
from returning true, which indicates it is closer to trigger the
action associated with the rule. For the example above,
\Fix{Lucas/Guilherme, it would be good to show how this looks like in
  code}
